---
title: "Retrieval-Augmented Generation"
subtitle: "Demo for deck 03: Prompt engineering and RAG"
format: html
---

Instructor demo: build a dynamic RAG system over the "R for Data Science" book.
We chunk the text, embed each chunk, and expose retrieval as a tool the chat
model can call.

Retrieval embeddings use OpenAI (Anthropic has no embeddings API), but the chat
model is Claude. Needs `OPENAI_API_KEY` and `ANTHROPIC_API_KEY` in your
`.Renviron`. Run the chunks interactively rather than rendering.

```{r}
#| label: setup
#| message: false
library(ragnar)
library(ellmer)
```

## Crawl, chunk, and embed the book

Crawl the R for Data Science website to find its pages, then create a `ragnar`
store that embeds each chunk with OpenAI's `text-embedding-3-small`. Reading a
page as markdown, chunking it, and inserting the chunks sends each chunk to
OpenAI for an embedding, which `ragnar` stores alongside the original text.

```{r}
base_url <- "https://r4ds.hadley.nz"
pages <- ragnar_find_links(base_url, children_only = TRUE)

store_location <- here::here("data/r4ds.ragnar.duckdb")

store <- ragnar_store_create(
  store_location,
  title = "R for Data Science",
  # Set overwrite = TRUE to rebuild from scratch.
  embed = \(x) embed_openai(x, model = "text-embedding-3-small")
)

for (page in pages) {
  chunks <- page |> read_as_markdown() |> markdown_chunk()
  ragnar_store_insert(store, chunks)
}

ragnar_store_build_index(store)
```

Or if you have already created the store:

```{r}
store <- ragnar_store_connect(store_location)
```

## Inspect what the store retrieves

Which chunks surface for a given question? The store inspector app lets you try
example questions against the store and see the retrieved chunks.

```{r}
ragnar_store_inspect(store)
```

## Chat with the book

`ragnar` provides a tool that attaches to an `ellmer` chat client and lets the
model retrieve relevant chunks on demand. Here the chat model is Claude; paste a
real "how do I ..." R question into the chat and watch it cite retrieved
passages.

```{r}
chat <- chat(
  "anthropic/claude-sonnet-4-6",
  system_prompt = r"--(
You are an expert R programmer and mentor. You are concise.

Before responding, retrieve relevant material from the knowledge store. Quote or
paraphrase passages, clearly marking your own words versus the source. Provide a
working link for every source you cite.
)--"
)

# Choose how many chunks to retrieve each time the model calls the tool.
ragnar_register_tool_retrieve(chat, store, top_k = 10)

live_browser(chat)
```
